---
title: "March Madness 2026: Prop Betting Analysis"
output: html_notebook
author: "Cat Murphy"
---
## Packages
```{r packages}
library(dplyr)
library(janitor)
library(tidyverse)
library(lubridate)
library(rvest)
library(stringr)
library(stringi)
library(readr)
library(purrr)
library(fs)
library(hms)
```

## Standardization Functions
### Standardize team names
```{r standardize_team_names}
# function to standardize team names in rotowire props and espn schedules
standardize_team_names <- function(team_name) {
  team_name <- str_remove_all(team_name, "@")
  team_name <- str_replace_all(team_name, "North Carolina State", "NC State")
  team_name <- str_replace_all(team_name, "UNCW", "UNC Wilmington")
  team_name <- str_replace_all(team_name, "St\\. Thomas \\(MN\\)", "St\\. Thomas-Minnesota")
  team_name <- str_replace_all(team_name, "Seattle", "Seattle U")
  team_name <- str_replace_all(team_name, "LIU", "Long Island University")
  team_name <- str_replace_all(team_name, "McNeese State", "McNeese")
  team_name <- str_replace_all(team_name, "\\bUNI\\b", "Northern Iowa")
  team_name <- str_replace_all(team_name, "Miami \\(FL\\)", "Miami")
  team_name <- str_replace_all(team_name, "Miami OH", "Miami \\(OH\\)")
  team_name <- str_replace_all(team_name, "\\bPenn\\b", "Pennsylvania")
  team_name <- str_replace_all(team_name, "Hawaii", "Hawai'i")
  team_name <- str_replace_all(team_name, "Prairie View$", "Prairie View A&M")
  team_name <- str_replace_all(team_name, "Queens NC", "Queens University")
  team_name <- str_replace_all(team_name, "Brigham Young", "BYU")
  team_name <- str_replace_all(team_name, "Saint Mary's \\(CA\\)", "Saint Mary's")
  team_name <- str_replace_all(team_name, "St\\. Mary's \\(CAL\\)", "Saint Mary's")
  team_name <- str_replace_all(team_name, "Saint Marys", "Saint Mary's")
  team_name <- str_replace_all(team_name, "Miami Ohio", "Miami \\(OH\\)")
  team_name <- str_replace_all(team_name, "Tennessee St\\.", "Tennessee State")
  team_name <- str_replace_all(team_name, "Wright St\\.", "Wright State")
  team_name <- str_replace_all(team_name, "Maryland-Baltimore County", "UMBC")
  team_name <- str_replace_all(team_name, "Connecticut", "UConn")
  team_name <- str_replace_all(team_name, "Long Island University-Brooklyn", "Long Island University")
  team_name <- str_replace_all(team_name, "Long Island$", "Long Island University")
  team_name <- str_replace_all(team_name, "Southern Methodist", "SMU")
  team_name <- str_replace_all(team_name, "Miami FL", "Miami")
  team_name <- str_replace_all(team_name, "Texas Christian", "TCU")
  team_name <- str_replace_all(team_name, "St. Johns", "St. John's")
  team_name <- str_replace_all(team_name, "Utah St\\.", "Utah State")
  team_name <- str_replace_all(team_name, "Central Florida", "UCF")
  team_name <- str_replace_all(team_name, "Virginia Commonwealth", "VCU")
  team_name <- str_replace_all(team_name, "Queens \\(NY\\)", "Queens University")
  return(team_name)
}
```

### Standardize team abbreviations
```{r standardize_team_abbreviations}
# function to standardize team abbreviations in espn spreads
match_school_abbreviations <- function(abbreviation) {
  case_when(
    abbreviation == "ALA" ~ "Alabama",
    abbreviation == "AKR" ~ "Akron",
    abbreviation == "ARIZ" ~ "Arizona",
    abbreviation == "ARK" ~ "Arkansas",
    abbreviation == "AUB" ~ "Auburn",
    abbreviation == "BRAD" ~ "Bradley",
    abbreviation == "BYU" ~ "BYU",
    abbreviation == "CAL" ~ "California",
    abbreviation == "CBU" ~ "California Baptist",
    abbreviation == "CLEM" ~ "Clemson",
    abbreviation == "CONN" ~ "UConn",
    abbreviation == "CSU" ~ "Colorado State",
    abbreviation == "DAY" ~ "Dayton",
    abbreviation == "DAV" ~ "Davidson",
    abbreviation == "DUKE" ~ "Duke",
    abbreviation == "FLA" ~ "Florida",
    abbreviation == "FUR" ~ "Furman",
    abbreviation == "GMU" ~ "George Mason",
    abbreviation == "GONZ" ~ "Gonzaga",
    abbreviation == "GW" ~ "George Washington",
    abbreviation == "HAW" ~ "Hawai'i",
    abbreviation == "HOF" ~ "Hofstra",
    abbreviation == "HOU" ~ "Houston",
    abbreviation == "HOW" ~ "Howard",
    abbreviation == "HP" ~ "High Point",
    abbreviation == "ILL" ~ "Illinois",
    abbreviation == "ILST" ~ "Illinois State",
    abbreviation == "IOWA" ~ "Iowa",
    abbreviation == "IDHO" ~ "Idaho",
    abbreviation == "ISU" ~ "Iowa State",
    abbreviation == "JOES" ~ "Saint Joseph's",
    abbreviation == "KENN" ~ "Kennesaw State",
    abbreviation == "KENT" ~ "Kent State",
    abbreviation == "KU" ~ "Kansas",
    abbreviation == "LEH" ~ "Lehigh",
    abbreviation == "LIB" ~ "Liberty",
    abbreviation == "LIU" ~ "Long Island University",
    abbreviation == "LOU" ~ "Louisville",
    abbreviation == "MCNS" ~ "McNeese",
    abbreviation == "MIA" ~ "Miami",
    abbreviation == "MICH" ~ "Michigan",
    abbreviation == "MIZZ" ~ "Missouri",
    abbreviation == "M-OH" ~ "Miami (OH)",
    abbreviation == "MSU" ~ "Michigan State",
    abbreviation == "MURR" ~ "Murray State",
    abbreviation == "NAVY" ~ "Navy",
    abbreviation == "NCST" ~ "NC State",
    abbreviation == "NDSU" ~ "North Dakota State",
    abbreviation == "NEB" ~ "Nebraska",
    abbreviation == "NEV" ~ "Nevada",
    abbreviation == "OKST" ~ "Oklahoma State",
    abbreviation == "OSU" ~ "Ohio State",
    abbreviation == "PENN" ~ "Pennsylvania",
    abbreviation == "PUR" ~ "Purdue",
    abbreviation == "PV" ~ "Prairie View A&M",
    abbreviation == "QUC" ~ "Queens University",
    abbreviation == "SCU" ~ "Santa Clara",
    abbreviation == "SEA" ~ "Seattle U",
    abbreviation == "SFA" ~ "Stephen F. Austin",
    abbreviation == "SHSU" ~ "Sam Houston",
    abbreviation == "SIE" ~ "Siena",
    abbreviation == "SJU" ~ "St. John's",
    abbreviation == "SLU" ~ "Saint Louis",
    abbreviation == "SMC" ~ "Saint Mary's",
    abbreviation == "SMU" ~ "SMU",
    abbreviation == "TCU" ~ "TCU",
    abbreviation == "TENN" ~ "Tennessee",
    abbreviation == "TEX" ~ "Texas",
    abbreviation == "TLSA" ~ "Tulsa",
    abbreviation == "TNST" ~ "Tennessee State",
    abbreviation == "TROY" ~ "Troy",
    abbreviation == "TTU" ~ "Texas Tech",
    abbreviation == "TXAM" ~ "Texas A&M",
    abbreviation == "UCF" ~ "UCF",
    abbreviation == "UCI" ~ "UC Irvine",
    abbreviation == "UCLA" ~ "UCLA",
    abbreviation == "UGA" ~ "Georgia",
    abbreviation == "UIC" ~ "UIC",
    abbreviation == "UK" ~ "Kentucky",
    abbreviation == "UMBC" ~ "UMBC",
    abbreviation == "UNC" ~ "North Carolina",
    abbreviation == "UNCW" ~ "UNC Wilmington",
    abbreviation == "UNI" ~ "Northern Iowa",
    abbreviation == "UNLV" ~ "UNLV",
    abbreviation == "UNM" ~ "New Mexico",
    abbreviation == "USA" ~ "South Alabama",
    abbreviation == "USF" ~ "South Florida",
    abbreviation == "UST" ~ "St. Thomas-Minnesota",
    abbreviation == "USU" ~ "Utah State",
    abbreviation == "UVA" ~ "Virginia",
    abbreviation == "UVU" ~ "Utah Valley",
    abbreviation == "VAN" ~ "Vanderbilt",
    abbreviation == "VCU" ~ "VCU",
    abbreviation == "VILL" ~ "Villanova",
    abbreviation == "WAKE" ~ "Wake Forest",
    abbreviation == "WICH" ~ "Wichita State",
    abbreviation == "WIS" ~ "Wisconsin",
    abbreviation == "WRST" ~ "Wright State",
    abbreviation == "WYO" ~ "Wyoming",
    abbreviation == "YALE" ~ "Yale")
}
```

### Standardize player names
```{r standardize_player_names}
standardize_player_names <- function(player_name) {
  player_name <- str_remove_all(player_name, ",? Jr\\.?")
  player_name <- str_remove_all(player_name, " II(I?)")
  player_name <- str_remove_all(player_name, "\\.")
  player_name <- stri_trans_general(player_name, "Latin-ASCII")
  player_name <- str_remove_all(player_name, "\".*?\"")
  player_name <- str_replace_all(player_name, "De Ridder", "De_Ridder")
  player_name <- str_replace_all(player_name, "Aleksandar Gavalyugov", "Sash Gavalyugov")
  player_name <- str_replace_all(player_name, "Solomon Ball", "Solo Ball")
  player_name <- str_replace_all(player_name, "Somtochukwu", "Somto")
  player_name <- str_replace_all(player_name, "Ben Vander Wal", "Ben VanderWal")
  player_name <- str_replace_all(player_name, "Timotej Malovec", "Timo Malovec")
  player_name <- str_replace_all(player_name, "Evan Williams", "Lance Williams")
  player_name <- str_squish(player_name)
  return(player_name)
}
```

### Standardize unique prop ID
```{r standardize_ids}
standardize_id <- function(prop_id) {
  prop_id <- str_replace_all(prop_id, "\\s+", "_")
  prop_id <- str_remove_all(prop_id, "\\.")
  prop_id <- str_remove_all(prop_id, "'")
  prop_id <- str_remove_all(prop_id, "-")
  prop_id <- str_remove_all(prop_id, ":")
  prop_id <- str_remove_all(prop_id, "\\(")
  prop_id <- str_remove_all(prop_id, "\\)")
  return(prop_id)
}
```

### Standardize prop market names
```{r standardize_market_names}
standardize_market_name <- function(market_name) {
  market_name <- str_remove_all(market_name, "Over/Under")
  market_name <- str_to_lower(market_name)
  market_name <- str_squish(market_name)
  market_name <- str_replace_all(market_name, "\\s?\\+\\s?", "_")
  market_name <- str_replace_all(market_name, "pts", "points")
  market_name <- str_replace_all(market_name, "3pt made", "three_pointers_made")
  market_name <- str_replace_all(market_name, "reb(?![A-Za-z])", "rebounds")
  market_name <- str_replace_all(market_name, "ast", "assists")
  market_name <- str_replace_all(market_name, "blk", "blocks")
  market_name <- str_replace_all(market_name, "stl", "steals")
  market_name <- str_replace_all(market_name, "-", "_")
  market_name <- str_replace_all(market_name, "\\s", "_")
  return(market_name)
}
```
### Classify platform types
```{r define_platform_type}
classify_platform_type <- function(platform_name) {
  case_when(
    platform_name == "betmgm" | platform_name == "thescore" | platform_name == "fanduel" | platform_name == "draftkings" | platform_name == "hardrock" | platform_name == "caesars" | platform_name == "sugarhouse" | platform_name == "betrivers" | platform_name == "partycasino" ~ "sportsbook",
    platform_name == "sleeper" | platform_name == "prizepicks" | platform_name == "underdog" | platform_name == "pick6" | platform_name == "fliff" ~ "dfs",
    TRUE ~ "other"
  )
}
```
### Fill in missing info
```{r fill_in_missing_info}
# write function to fill in missing game info  
missing_game_info <- function(scrape_date, player_name) {
 case_when(
   scrape_date == "2026-03-19" & player_name == "Kennard Davis" ~ "BYU at TEX",
   scrape_date == "2026-03-19" & player_name == "AJ Dybantsa" ~ "BYU at TEX",
   scrape_date == "2026-03-19" & player_name == "AK Okereke" ~ "MCNS at VAN",
   scrape_date == "2026-03-19" & player_name == "CJ Brown" ~ "USF at LOU",
   scrape_date == "2026-03-19" & player_name == "RJ Johnson" ~ "KENN at GONZ",
   scrape_date == "2026-03-19" & player_name == "AJ Levine" ~ "PENN at ILL",
   scrape_date == "2026-03-19" & player_name == "TJ Power" ~ "PENN at ILL",
   scrape_date == "2026-03-20" & player_name == "TJ Burch" ~ "WRST at UVA",
   scrape_date == "2026-03-20" & player_name == "MJ Collins" ~ "USU at VILL",
   scrape_date == "2026-03-21" & player_name == "AK Okereke" ~ "VAN at NEB",
   scrape_date == "2026-03-22" & player_name == "MJ Collins" ~ "USU at ARIZ")
}
```

## Load Data
### Load Rosters
```{r load_rosters}
roster_paths <- dir_ls("../data/rosters/teams", recurse = TRUE, regexp = "\\.csv")
  
rosters <- roster_paths |>
  map_dfr(read_csv, show_col_types = FALSE) |>
  select(1:2) |>
  rename(player_team = team) |>
  mutate(player_name = standardize_player_names(player_name),
         player_team = standardize_team_names(player_team))
```
### Load ESPN Schedules
```{r load_espn_schedules}
# load ESPN file paths
schedules_paths <- dir_ls("../data/espn_schedules", recurse = TRUE, regexp = "\\.csv")

# load and clean schedules
schedules <- schedules_paths |>
  map_dfr(read_csv, show_col_types = FALSE) |>
  clean_names() |>
  
  separate(game_date, into = c("game_day", "game_date", "game_year"), sep = ", ") |>
  
  mutate(game_date = mdy(paste(game_date, game_year, sep = "-"))) |>

  separate(venue, into = c("venue", "city", "state"), sep = ", ") |>
  
  mutate(game = paste(pmin(home_team, away_team), "vs.", pmax(home_team, away_team)),
         favored_team = match_school_abbreviations(spread_team)
         ) |>
  
  rename(game_points_spread = spread_line) |>
  
  select(game, game_date, game_time, tournament, round, region, away_team, away_team_seed, home_team, home_team_seed, venue, city, state, tv_network, game_points_spread, favored_team) |>
  
  distinct(game, .keep_all = TRUE)
```
### Load Box Score Data
```{r}
box_score_paths <- dir_ls("../data/box_scores", recurse = TRUE, regexp = "\\.csv")

box_scores <- box_score_paths |>
  map_dfr(read_csv, show_col_types = FALSE) |>
  clean_names() |>
  rename(player_team = team_location,
         opponent_team = opponent_location) |>
  separate(date, into = c("game_date", "game_time"), sep = " ") |>
  mutate(game_date = as.Date(game_date),
         player_name = standardize_player_names(player_name),
         game = paste(pmin(player_team, opponent_team), "vs.", pmax(player_team, opponent_team)))  |>
  select(game, game_date, player_team, opponent_team, player_name, minutes) |>
  rename(player_minutes = minutes)
```
### Load Rotowire Props
#### Load Subscription Rotowire props
```{r subscription_rotowire_props}
subscription_rotowire_props_paths <- dir_ls("../data/rotowire_subscription/lines", recurse = TRUE, regexp = "\\.csv")

subscription_rotowire_props <-  subscription_rotowire_props_paths |>
  map_dfr(~ read_csv(.x, show_col_types = FALSE) |> mutate(file_name = .x)) |>
  clean_names() |>
  distinct() |>
  mutate(row_id = row_number(),
         data_source = "rotowire_subscription",
         file_date = as.Date(str_extract(file_name, "\\d{4}-\\d{2}-\\d{2}"))) |>
  select(row_id, data_source, file_date, player, team, opponent, market_name, site, line, file_name) |>
  rename(player_name = player,
         player_team = team,
         opponent_team = opponent,
         prop_platform = site,
         prop_market = market_name,
         prop_line = line) |>
  mutate(prop_platform = str_remove_all(prop_platform, "-sb"),
         across(c(player_team, opponent_team), \(x) str_remove_all(x, "@")),
         prop_market = standardize_market_name(prop_market),
         prop_platform_type = classify_platform_type(prop_platform),
         player_name = standardize_player_names(player_name),
         player_team = standardize_team_names(player_team),
         opponent_team = standardize_team_names(opponent_team),
         game = paste(pmin(player_team, opponent_team), "vs.", pmax(player_team, opponent_team))) |>
  
  left_join(schedules, join_by(game)) |>
  
  filter(tournament == "NCAA Tournament") |>
  
  mutate(player_team_seed = case_when(
    player_team == home_team ~ home_team_seed,
    player_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  opponent_team_seed = case_when(
    opponent_team == home_team ~ home_team_seed,
    opponent_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  player_team_favored = case_when(
    player_team == favored_team ~ TRUE,
    opponent_team == favored_team ~ FALSE,
    TRUE ~ NA
  )) |>
  
  mutate(prop_id = str_squish(str_to_lower(paste(player_name, prop_platform, prop_market, "no", player_team_seed, player_team, "vs", "no", opponent_team_seed, opponent_team, game_date, sep = "_"))),
         prop_id = standardize_id(prop_id)) |>

  select(
    row_id,
    data_source,
    file_name,
    file_date,
    tournament, 
    round, 
    region, 
    game,
    game_date,
    player_name, 
    player_team, 
    player_team_seed, 
    opponent_team, 
    opponent_team_seed, 
    game_points_spread, 
    favored_team,
    player_team_favored,
    prop_platform, 
    prop_platform_type,
    prop_market, 
    prop_line,
    prop_id)
```

#### Load Scraped Rotowire props
```{r scraped_rotowire_props}
# load file paths
scraped_rotowire_props_paths <- dir_ls("../data/rotowire_props", recurse = TRUE, regexp = "\\.csv")

scraped_rotowire_props <- scraped_rotowire_props_paths |>
  map_dfr(~ read_csv(.x, show_col_types = FALSE) |> mutate(file_name = .x)) |>
  clean_names() |>
  select(-scrape_date, -scrape_time) |>
  mutate(row_id = row_number(),
         data_source = "scraped_rotowire",
         file_date = as.Date(str_extract(file_name, "\\d{4}-\\d{2}-\\d{2}"))) |>
  rename(player_team = team,
         opponent_team = opponent,
         prop_market = stat_type,
         betmgm_line = mgm_line) |>
  mutate(prop_market = standardize_market_name(prop_market),
         player_name = standardize_player_names(player_name),
         across(c(player_team, opponent_team), \(x) standardize_team_names(x)),
         game = paste(pmin(player_team, opponent_team), "vs.", pmax(player_team, opponent_team))) |>
  left_join(schedules, join_by(game)) |>
  
  filter(tournament == "NCAA Tournament") |>
  
  mutate(player_team_seed = case_when(
    player_team == home_team ~ home_team_seed,
    player_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  opponent_team_seed = case_when(
    opponent_team == home_team ~ home_team_seed,
    opponent_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  player_team_favored = case_when(
    player_team == favored_team ~ TRUE,
    opponent_team == favored_team ~ FALSE,
    TRUE ~ NA
  )) |>

  pivot_longer(
    cols = ends_with("_line"),
    names_to = "prop_platform",
    values_to = "prop_line",
    names_transform = \(x) str_remove(x, "_line")) |>
  
  filter(!is.na(prop_line)) |>
  mutate(prop_platform_type = classify_platform_type(prop_platform),
         prop_id = str_squish(str_to_lower(paste(player_name, prop_platform, prop_market, "no", player_team_seed, player_team, "vs", "no", opponent_team_seed, opponent_team, game_date, sep = "_"))),
         prop_id = standardize_id(prop_id)) |>

  select(
    row_id,
    data_source,
    file_name,
    file_date,
    tournament, 
    round, 
    region, 
    game,
    game_date,
    player_name, 
    player_team, 
    player_team_seed, 
    opponent_team, 
    opponent_team_seed, 
    game_points_spread, 
    favored_team,
    player_team_favored,
    prop_platform, 
    prop_platform_type,
    prop_market, 
    prop_line,
    prop_id) 
```


### Load DraftKings props
#### Load DraftKings Top Props
```{r draftkings_top_props}
draftkings_top_50_paths <- dir_ls("../data/draftkings_props/top_50", recurse = TRUE, regexp = "\\.csv")

draftkings_top_props <- draftkings_top_50_paths |>
  map_dfr(~ read_csv(.x, show_col_types = FALSE) |> mutate(file_name = .x)) |>
  clean_names() |>
  
  separate(raw_data, into = c("game", "game_date", "prop", "prop_line", "odds"), sep = " \\| ", extra = "merge") |>
  separate(game, into = c("team1", "team2"), sep = " @ ") |>
  separate(game_date, into = c("game_date", "game_time"), sep = ", ") |>
  
  mutate(
    row_id = row_number(),
    data_source = "draftkings_top_props",
    file_date = as.Date(str_extract(file_name, "\\d{4}-\\d{2}-\\d{2}")),
    
    across(c(team1, team2), \(x) standardize_team_names(x)),
    
    game = paste(pmin(team1, team2), "vs.", pmax(team1, team2)),
    game_date = as.Date(game_date, format = "%m/%d"),
    
    prop = case_when(
      prop == "First Points Scorer" ~ paste(prop_line, prop, sep = " "),
      TRUE ~ prop
  ),
    prop = standardize_player_names(prop),
        
    prop_line = str_remove_all(prop_line, "\\D"),
    prop_line = as.numeric(prop_line)
  ) |>
  
  separate(prop, into = c("first_name", "last_name", "prop_market"), sep = " ", extra = "merge") |>
  
  mutate(
    player_name = paste(first_name, last_name),
    prop_market = standardize_market_name(prop_market),
    prop_platform = "draftkings",
    prop_platform_type = classify_platform_type(prop_platform)) |>
  
  select(row_id, data_source, file_name, file_date, player_name, prop_platform, prop_market, prop_platform_type, prop_line, team1, team2, game, game_date) |>

  left_join(rosters, join_by(player_name)) |> 
  left_join(schedules, join_by(game, game_date)) |>
  
  filter(tournament == "NCAA Tournament") |>
  
  mutate(player_team_seed = case_when(
    player_team == home_team ~ home_team_seed,
    player_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  
  opponent_team = case_when(
    player_team == home_team ~ away_team,
    player_team == away_team ~ home_team,
    TRUE ~ NA
  ),
  
  opponent_team_seed = case_when(
    opponent_team == home_team ~ home_team_seed,
    opponent_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  
  player_team_favored = case_when(
    player_team == favored_team ~ TRUE,
    opponent_team == favored_team ~ FALSE,
    TRUE ~ NA)) |>
  
  mutate(prop_id = str_squish(str_to_lower(paste(player_name, prop_platform, prop_market, "no", player_team_seed, player_team, "vs", "no", opponent_team_seed, opponent_team, game_date, sep = "_"))),
         prop_id = standardize_id(prop_id)) |>

  select(
    row_id,
    data_source,
    file_name,
    file_date,
    tournament, 
    round, 
    region, 
    game,
    game_date,
    player_name, 
    player_team, 
    player_team_seed, 
    opponent_team, 
    opponent_team_seed, 
    game_points_spread,
    favored_team,
    player_team_favored,
    prop_platform,
    prop_platform_type,
    prop_market,
    prop_line,
    prop_id)
```
#### Load DraftKings Sorted Props
```{r draftkings_sorted_props}
draftkings_sorted_props_paths <- dir_ls("../data/draftkings_props/sorted_props", recurse = TRUE, regexp = "\\.csv")

draftkings_sorted_props <- draftkings_sorted_props_paths |>
  map_dfr(~ read_csv(.x, show_col_types = FALSE) |> mutate(file_name = .x)) |>
  select(file_name, scrape_date, matchup, prop_type, player_name) |>
  clean_names() |>
  distinct() |>
  filter(matchup != "Unknown") |>
  mutate(row_id = row_number(),
         data_source = "draftkings_sorted_props",
         file_date = as.Date(str_extract(file_name, "\\d{4}-\\d{2}-\\d{2}"))) |>
  separate(matchup, into = c("team1", "team2"), sep = " @ ") |>
  
  mutate(
    across(c(team1, team2), \(x) standardize_team_names(x)),
    
    game = paste(pmin(team1, team2), "vs.", pmax(team1, team2)),
    
    prop = case_when(
      prop_type == "First Points Scorer" | prop_type == "Assists" | prop_type == "Points"| prop_type == "Points + Rebounds + Assists"| prop_type == "Points + Rebounds" | prop_type == "Points + Assists" | prop_type == "Rebounds" | prop_type == "Three Pointers Made" ~ paste(player_name, prop_type),
      TRUE ~ prop_type
    ),
  
    prop = standardize_player_names(prop)
  ) |>
  
  separate(prop, into = c("first_name", "last_name", "prop_market"), sep = " ", extra = "merge") |>
  
  mutate(
    player_name = paste(first_name, last_name),
    prop_market = standardize_market_name(prop_market),
    prop_platform = "draftkings",
    prop_platform_type = classify_platform_type(prop_platform),
    prop_line = NA
  ) |>
  
  select(row_id, data_source, file_name, file_date, game, player_name, prop_platform, prop_market, prop_platform_type, prop_line) |>
  
  left_join(rosters, join_by(player_name)) |>
  filter(!is.na(player_team)) |>
  left_join(schedules, join_by(game)) |>
  
  filter(tournament == "NCAA Tournament") |>
  
  mutate(player_team_seed = case_when(
    player_team == home_team ~ home_team_seed,
    player_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  
  opponent_team = case_when(
    player_team == home_team ~ away_team,
    player_team == away_team ~ home_team,
    TRUE ~ NA
  ),
  
  opponent_team_seed = case_when(
    opponent_team == home_team ~ home_team_seed,
    opponent_team == away_team ~ away_team_seed,
    TRUE ~ NA
  ),
  
  player_team_favored = case_when(
    player_team == favored_team ~ TRUE,
    opponent_team == favored_team ~ FALSE,
    TRUE ~ NA)) |>
  
  mutate(prop_id = str_squish(str_to_lower(paste(player_name, prop_platform, prop_market, "no", player_team_seed, player_team, "vs", "no", opponent_team_seed, opponent_team, game_date, sep = "_"))),
         prop_id = standardize_id(prop_id)) |>
  
  select(
    row_id,
    data_source,
    file_name,
    file_date,
    tournament, 
    round, 
    region, 
    game,
    game_date,
    player_name, 
    player_team, 
    player_team_seed, 
    opponent_team, 
    opponent_team_seed, 
    game_points_spread,
    favored_team,
    player_team_favored,
    prop_platform,
    prop_platform_type,
    prop_market, 
    prop_line,
    prop_id)
```
### Load Bettingpros Props
```{r bettingpros_props}
bettingpros_props_paths <- dir_ls("../data/bettingpros_props", recurse = TRUE, regexp = "\\.csv")

bettingpros_props <- bettingpros_props_paths |>
  map_dfr(~ read_csv(.x, show_col_types = FALSE) |> mutate(file_name = .x)) |>
  clean_names() |>
  rename(prop_market = market_type,
         thescore_line = thescore_bet_line,
         prophetx_line = prophetx_light_line) |>
  select(scrape_date, scrape_time, file_name, player_name, game_matchup, prop_market, ends_with("_line")) |>

  mutate(
    row_id = row_number(),
    data_source = "bettingpros",
    file_date = as.Date(str_extract(file_name, "\\d{4}-\\d{2}-\\d{2}")),
    player_name = standardize_player_names(player_name),
    prop_market = standardize_market_name(prop_market),
  
    missing_game_info = missing_game_info(scrape_date, player_name),
    
    game_matchup = case_when(
        !is.na(missing_game_info) ~ missing_game_info,
        TRUE ~ game_matchup
  )) |>
  # filter(!is.na(game_matchup)) |>
  separate(game_matchup, into = c("team1", "team2"), sep = " at ") |>
  mutate(across(c(team1, team2), \(x) match_school_abbreviations(x)),
         game = str_squish(paste(pmin(team1, team2), "vs.", pmax(team1, team2)))) |>
  left_join(schedules, join_by(game)) |>
  left_join(rosters, join_by(player_name)) |>
  filter(tournament == "NCAA Tournament") |>
  mutate(
    player_team_seed = case_when(
      player_team == home_team ~ home_team_seed,
      player_team == away_team ~ away_team_seed,
      TRUE ~ NA
    ),
    opponent_team = case_when(
      player_team == home_team ~ away_team,
      player_team == away_team ~ home_team,
      TRUE ~ NA
    ),
    opponent_team_seed = case_when(
      opponent_team == home_team ~ home_team_seed,
      opponent_team == away_team ~ away_team_seed,
      TRUE ~ NA
    ),
    player_team_favored = case_when(
      player_team == favored_team ~ TRUE,
      opponent_team == favored_team ~ FALSE,
      TRUE ~ NA
    )) |>
  pivot_longer(
    cols = ends_with("line"), 
    names_to = "prop_platform", 
    values_to = "prop_line", 
    names_transform = \(x) str_remove(x, "_line")) |>
  
  filter(!is.na(prop_line)) |>
  mutate(prop_platform_type = classify_platform_type(prop_platform),
         prop_id = str_squish(str_to_lower(paste(player_name, prop_platform, prop_market, "no", player_team_seed, player_team, "vs", "no", opponent_team_seed, opponent_team, game_date, sep = "_"))),
         prop_id = standardize_id(prop_id)) |>
  
  select(
    row_id,
    data_source,
    file_name,
    file_date,
    tournament, 
    round, 
    region, 
    game,
    game_date,
    player_name, 
    player_team, 
    player_team_seed, 
    opponent_team, 
    opponent_team_seed, 
    game_points_spread,
    favored_team,
    player_team_favored,
    prop_platform, 
    prop_platform_type,
    prop_market, 
    prop_line,
    prop_id)
```

### Bind All Rows
```{r}
all_props <- bind_rows(subscription_rotowire_props, scraped_rotowire_props, draftkings_top_props, draftkings_sorted_props, bettingpros_props) |>
  group_by(prop_id) |>
  mutate(number_times_identified = n(),
         file_date = paste(unique(file_date), collapse = ", "),
         data_source = paste(unique(data_source), collapse = ", "),
         file_name = paste(unique(file_name), collapse = ", ")) |>
  arrange(prop_id, desc(file_date)) |>
  slice_head(n = 1) |>
  ungroup() |>
  left_join(box_scores, join_by(game, game_date, player_name, player_team, opponent_team)) |>
  select(
    tournament, 
    round, 
    game,
    game_date,
    player_name, 
    player_team, 
    player_team_seed, 
    player_minutes,
    opponent_team, 
    opponent_team_seed, 
    game_points_spread,
    favored_team,
    player_team_favored,
    prop_platform, 
    prop_platform_type,
    prop_market, 
    prop_line,
    prop_id,
    number_times_identified, 
    data_source, 
    file_date, 
    file_name)

remove(subscription_rotowire_props, scraped_rotowire_props, draftkings_top_props, draftkings_sorted_props, bettingpros_props, box_scores, schedules, rosters)

remove(standardize_team_names, match_school_abbreviations, standardize_player_names, standardize_id, standardize_market_name, classify_platform_type, missing_game_info)

remove(draftkings_sorted_props_paths, draftkings_top_50_paths, scraped_rotowire_props_paths, subscription_rotowire_props_paths, bettingpros_props_paths, box_score_paths, schedules_paths, roster_paths)

write_csv(all_props |> select(-file_name), "../data/cleaned_props/cleaned_props.csv")
```



